FTLR proof cases

### 2018-06-06 fully-abstract-comp/ftlr-proof-cases.tex
FTLR proof changes

Let $n' \leq n$ be given and assume
\begin{enumerate}
\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$ \label{item:reg-ass}
\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
\end{enumerate}
Further let
\item $\Phi_S = (\ms_S,\reg_S\update{\pcreg}{c},\stk,\ms_\stk)$
\item $\Phi_T = (\ms_T,\reg_T\update{\pcreg}{c})$
and show
  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
In order to do this, we first, we prove that one of the following holds:
\lau{TODO: add $\neq$ failed to the cases that are not failed.}
  \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T
    \step[i] \failed$ for some $i$.
  \item \label{case:ftlr:halted} $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
  \item \label{case:ftlr:get-and-arith-op} All of the following hold: (includes simple cases: gettype, geta, getb,
    gete, getp, getl, lt, plus, minus, one case of move that can be handled
    \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r}{z}} \neq \failed$
    \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r}{z}} \neq \failed$
    \item $z \in \ints$
  \item \label{case:ftlr:call} All of the following hold:
    \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
    \item $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$
    \item $\exec{\Phi_S(\pcreg)}$
    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in
      \ints$.
    \item $\Phi_S \step[\gc] \Phi_S' \neq \failed$
    \item $\Phi_T \step \Phi_T' \neq \failed$
    \item $\Phi_T' = \updPc{\Phi_T\updReg{\rtmp{1}}{z}}$
    \item $z \in \ints$
    \item $\Phi_S(r_1) = \sealed{\sigma,c_1}$
    \item $\Phi_S(r_2) = \sealed{\sigma,c_2}$
    \item $\nonExec{c_2}$
    \item $\Phi_S(\stk) = \stkptr{\rw, \baddr_\stk, \eaddr_\stk, \aaddr_\stk}$
    \item $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$
    \item $\Phi_S(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
    \item $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$
    \item
          \Phi_S'.\reg = \Phi_S.\reg
          &[\pcreg \mapsto c_1] \\
          &[\rdata \mapsto c_2] \\
          &[\rstk \mapsto \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}] \\
          &[\rretc \mapsto \sealed{\sigma',\retptrc{\baddr,\eaddr,\aaddr+\calllen}}]\\
          &[\rretd \mapsto \sealed{\sigma',\retptrd{\aaddr_\stk,\eaddr_\stk}}] \\
          &[r_1,r_2 \mapsto w_1,w_2] \\
          &[\rtmp{1} \mapsto 0]
    \item $\Phi_S'.\mem = \Phi_S.\mem$
    \item $\Phi_S'.\stk = ((\aaddr+\calllen),\ms_\stk |_{\aaddr_\stk,\eaddr_\stk}\update{\aaddr_\stk}{42}) :: \Phi_S.\stk$
    \item $\Phi_S'.\ms_\stk = \Phi_S.\ms_\stk = \ms_\stk - \ms_\stk|_{\aaddr_\stk,\eaddr_\stk}$
    \item $\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}$
    \item $\sigma' = \sigma_\aaddr + \offsigma$
    \item $\sigma_\aaddr \leq \sigma' \leq \sigma_\eaddr$

@@

In case \ref{case:ftlr:failed}, both executions go to failed. In this case, we
either assume that the target configuration terminates or that the source
configuration terminates, but in either case it leads to a contradiction with
the failing configuration.
\\\\
In case \ref{case:ftlr:halted}, both source and target configuration halts in 0
steps, so both directions of $\lro$ are trivially satisfied.
\\\\
In case \ref{case:ftlr:get-and-arith-op}, we use the induction hypothesis to get
\[
  \npair[n'-1]{(((\perm,\lin),\baddr,\eaddr,\aaddr+1),((\perm,\lin),\baddr,\eaddr,\aaddr+1))} \in \lre(W)
\]
This is possible by using the initial assumptions along with Lemma \ref{lem:downwards-closed}.
By Lemma~\ref{lem:downwards-closed}, we know $\memSat[n'-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
By Lemma~\ref{lem:downwards-closed} and the fact that integers are always safe,
we know $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$.
This allows us to conclude
\[
  \npair[n'-1]{(\Phi_S',\Phi_T')} \in \lro
\]
as desired.

>>>>

Let $n' \leq n$ be given and assume
\begin{enumerate}
\item $\npair[n']{\stpair{\reg}{\reg}} \in \lrrg{\trust}(W_R)$ \label{item:reg-ass}
\item $\memSat[n']{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$\label{item:mem-ass}
\item $W \oplus W_R \oplus W_M$ is well-defined\label{item:worlds-def}
\end{enumerate}
Further let
\begin{itemize}
\item $\Phi_S = (\ms_S,\reg_S\update{\pcreg}{c},\stk,\ms_\stk)$
\item $\Phi_T = (\ms_T,\reg_T\update{\pcreg}{c})$
\end{itemize}
and show
\[
  \npair[n']{(\Phi_S,\Phi_T)} \in \lro
\]
In order to do this, we first, we prove that one of the following holds:
\lau{TODO: add $\neq$ failed to the cases that are not failed.}
\begin{enumproof}
  \item \label{case:ftlr:failed} $\Phi_S \step[\gc] \failed$ and $\Phi_T
    \step[i] \failed$ for some $i$.
  \item \label{case:ftlr:halted} $\Phi_S \step[\gc] \halted$ and $\Phi_T \step\halted$
  \item \label{case:ftlr:get-and-arith-op} All of the following hold: (includes simple cases: gettype, geta, getb,
    gete, getp, getl, lt, plus, minus, one case of move that can be handled
    uniformly)
    \begin{itemize}
    \item $\Phi_S \step[\gc] \Phi_S'$
    \item $\Phi_T \step \Phi_T'$
    \item $\Phi_S' = \updPcAddr{\Phi_S\updReg{r}{z}} \neq \failed$
    \item $\Phi_T' = \updPcAddr{\Phi_T\updReg{r}{z}} \neq \failed$
    \item $z \in \ints$
    \end{itemize}
  \item \label{case:ftlr:call} All of the following hold:
    \begin{itemize}
    \item $\callCond{\Phi_S,r_1,r_2,\aaddr}$
    \item $\baddr \leq \aaddr \tand \aaddr + \calllen-1 \leq \eaddr$
    \item $\exec{\Phi_S(\pcreg)}$
    \item for all $i = 0..\calllen-1$, $\ms_T(\aaddr+i) = \ms_S(\aaddr+i) \in \ints$.
    \item $[\baddr,\eaddr] \subseteq \ta$
    \item $\Phi_S \step[\gc] \Phi_S' \neq \failed$
    \item $\Phi_T \step \Phi_T' \neq \failed$
    \item $\Phi_T' = \updPc{\Phi_T\updReg{\rtmp{1}}{z}}$
    \item $z \in \ints$
    \item $\Phi_S(r_1) = \sealed{\sigma,c_1}$
    \item $\Phi_S(r_2) = \sealed{\sigma,c_2}$
    \item $\nonExec{c_2}$
    \item $\Phi_S(\stk) = \stkptr{\rw, \baddr_\stk, \eaddr_\stk, \aaddr_\stk}$
    \item $\baddr_\stk < \aaddr_\stk \leq \eaddr_\stk$
    \item $\Phi_S(\pcreg) = ((\perm,\lin),\baddr,\eaddr,\aaddr)$
    \item $w_1 = \linCons{c_1}$ and $w_2 = \linCons{c_2}$
    \item \[
        \begin{split}
          \Phi_S''.\reg = \Phi_S.reg      &[\rstk \mapsto \stkptr{\rw,\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1}] \\
          &[\rretc \mapsto \sealed{\sigma',\retptrc(\baddr,\eaddr,\aaddr+\calllen)}]\\
          &[\rretd \mapsto \sealed{\sigma',\retptrd(\aaddr_\stk,\eaddr_\stk)}] \\
          &[r_1,r_2 \mapsto w_1,w_2] \\
          &[\rtmp{1} \mapsto 0]
        \end{split}
      \]
    \item $\Phi_S''.\mem = \Phi_S.\mem$
    \item $\ms_{\stk\_\priv,S} = \Phi_S.\ms_\stk |_{\aaddr_\stk,\eaddr_\stk}\update{\aaddr_\stk}{42}$
    \item $\Phi_S''.\stk = ((\aaddr+\calllen),\ms_{\stk\_\priv,S}) :: \Phi_S.\stk$
    \item $\Phi_S''.\ms_\stk = \Phi_S.\ms_\stk -
      \Phi_S.\ms_\stk|_{\aaddr_\stk,\eaddr_\stk}$
    \item $\baddr \leq \aaddr+\offpc \leq \eaddr$
      \lau{TODO: Need to have an explicit $\offpc$.}
    \item $\mem(\aaddr+\offpc) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma_\aaddr}$
    \item $\sigma' = \sigma_\aaddr + \offsigma$
    \item $\sigma_\aaddr \leq \sigma' \leq \sigma_\eaddr$

    \item $\Phi_S' =\var{xjumpResult}\left(c_1,c_2, \Phi_S' \right)$

@@

In case \ref{case:ftlr:failed}, both executions go to failed. In this case, we
either assume that the target configuration terminates or that the source
configuration terminates, but in either case it leads to a contradiction with
the failing configuration.
\\\\
In case \ref{case:ftlr:halted}, both source and target configuration halts in 0
steps, so both directions of $\lro$ are trivially satisfied.
\\\\
In case \ref{case:ftlr:get-and-arith-op}, we use the induction hypothesis to get
\[
  \npair[n'-1]{(((\perm,\lin),\baddr,\eaddr,\aaddr+1),((\perm,\lin),\baddr,\eaddr,\aaddr+1))} \in \lre(W)
\]
This is possible by using the initial assumptions along with Lemma \ref{lem:downwards-closed}.
By Lemma~\ref{lem:downwards-closed}, we know $\memSat[n'-1]{\ms_S,\stk,\ms_\stk,\ms_T}{W_M}$.
By Lemma~\ref{lem:downwards-closed} and the fact that integers are always safe,
we know $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$.
This allows us to conclude
\[
  \npair[n'-1]{(\Phi_S',\Phi_T')} \in \lro
\]
as desired.
\\\\
In case \ref{case:ftlr:call}, we first need to let the target execution catch
up. That is $\Phi_T' \nstep[15]{} \Phi_T''$ for 
\[
\Phi_T'' = \Phi_T
\begin{array}[t]{l}
  \update{\mem.\aaddr_\stk}{42} \\
  \updReg{\rstk}{((\rw,\linear),\baddr_\stk,\aaddr_\stk-1,\aaddr_\stk-1)} \\
  \updReg{\rretd}{\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk-1)}} \\
  \updReg{\rretc}{\sealed{\sigma',((\perm,\lin),\baddr,\eaddr,\aaddr+\var{off\_return})}} \\
  \updReg{\rtmp{1}}{0}\\
  \updReg{r_1,r_2}{\linCons{\Phi_T(r_1)},\linCons{\Phi_T(r_2)}}\\
  \updReg{\pcreg}{c_1'}\\
  \updReg{\rdata}{c_2'}
\end{array}
\]
where $\Phi_T(r_i) = \sealed{\sigma,c_i'} \text{ for $i\in\{1,2\}$}$ and $\var{off\_return} = 15$ which is the offset to the return code.
Now using Lemma~\ref{lem:lro-anti-red-gen} again, it suffices to show
\[
  \npair[n'-1]{(\Phi_S',\Phi_T'')} \in \lro
\]
By assumption, we have $\npair[n']{(\reg_S(r_i),\reg_T(r_i))} \in \lrr(W_{R,i})$ for some $W_{R,i}$ with $i \in \{1,2\}$.
We know the capabilities in $r_1$ and $r_2$ are sealed capabilities, and by
Lemma~\ref{lem:unique-h-sigma} and the definition of $\lrv$ we get
$\npair[n'-1]{\stpair[.]{c_2}{c_2'}} \in H_\sigma \; \sigma \; \xi^{-1}(W_{R,2})$
and w.l.o.g.\

\begin{equation}
\forall W' \future W_{R,1}, W_o, n' < n, \npair[n']{(c_2,c_2')} \in H_\sigma \; \sigma \; \xi^{-1}(W_o) \ldotp \npair[n']{c_1,c_2,c_1',c_2'} \in \lrexj(W'\oplus W_o)\label{eq:ftlr:call:sealed-code-assump}
\end{equation}
Now construct $W_{R,1}'$ as follows:

By Lemma~\ref{lem:nsub-and-nsup-std} and the safety assumption on the register-file, there exists $S \supseteq [\baddr_\stk, \eaddr_\stk]$ such that for some $R : S \fun \powerset{\nats}$ we have $\biguplus_{r \in S} R(r) \supseteq [\baddr_\stk,\eaddr_\stk]$ and for all $r$ in $S$ $\pwfree[W_{R,1}](r) \nsim \stdreg{R(r),\gc}{\pur}$ and $|R(r)| = 1$ and $\pwfree[W_{R,1}](r) \text{ is address-stratified}$.
Now define
\[
  W_{R,1}' = W_{R,1}[\free.R^{-1}([\aaddr_\stk,\eaddr_\stk]) \mapsto \revoked][\priv.r_\var{priv\_stk} \mapsto \stareg[(\ms_{\stk\_\priv,S},\Phi_T''.\mem|_{[\aaddr_\stk,\eaddr_\stk]}),\gc]{\spa,\square}]
\] 
For a fresh $r_\var{priv\_stk}$.
We know $W_{R,1}' \future W_{R,1}$ as the revoked regions must have been $\spatial$ in $W_{R,q}$ (as they are owned by the part of the world assigned to the stack-register in the register-file relation).
The static region for the private stack is an extension.

Pick this world as $W'$ in Eq~\ref{eq:ftlr:call:sealed-code-assump}
. Let $W_{R,2}'$ be the same world but
with the ownership of $W_{R,2}$ and pick it for $W_o$.
Now use monotonicity of $H_\sigma$ with the above facts to get
\[
  \npair[n']{c_1,c_2,c_1',c_2'} \in \lrexj(W_{R,1}'\oplus W_{R,2}')
\]
Now pick register files and memories such that they form $\Phi_S''$ (defined in
the assumptions) and $\Phi_T''$ and for $W_R'$ and $W_M'$ (we define them below) show
\begin{enumerate}[label=\roman*.]
\item \label{case:ftlr:scall:world} $W_{R,1}'\oplus W_{R,2}' \oplus W_R' \oplus W_M'$
\item \label{case:ftlr:scall:reg} $\npair[n'-1]{(\Phi_S''.\reg,\Phi_T''.\reg)} \in \lrr(\{\rdata\})(W_R')$
\item \label{case:ftlr:scall:mem} $\memSat[n'-1]{\Phi_S''.\mem,\Phi_S''.\stk,\Phi_S''.\ms_\stk,\Phi_T''.\mem}{W_M'}$
\end{enumerate}
to get
\[
 \npair[n'-1]{(\Phi_S',\Phi_T'')} \in \lro
\]
as desired. It remains to show \ref{case:ftlr:scall:world}-\ref{case:ftlr:scall:mem}.

Case \ref{case:ftlr:scall:world}: Pick $W_R'$ and $W_M'$ to have the regions of $W_{R,1}'$, but where $W_R'$ owns $r_\var{priv\_stk}$ and otherwise has the ownership of $W_R$ and $W$ with the exception of the regions owned by $W_{R,1}'$ and $W_{R,2}'$.
$W_M'$ has the ownership of $W_M$.
Case \ref{case:ftlr:scall:world} follows from assumption $W \oplus W_R \oplus W_M$.
The only changes to the worlds is that some ownership has been shifted from $W_R$ to $W_{R,1}$ and $W_{R,2}$ and the ownership for $W$ now belongs to $W_R'$.
In other words, no ownership has been duplicated.

Case \ref{case:ftlr:scall:reg}, using assumption $\npair[n'-1]{(\reg_S\update{r}{z},\reg_T\update{r}{z})} \in \lrr(W_R)$, we get a way to split the ownership of $W_R$.
We take this as the starting point, but with the following changes: regions $r_1$ and $r_2$ maps to worlds with no ownership.
region $\rstk$ maps to a world with the same ownership, but of course without the now revoked regions.
Region $\rretd$ maps to a world that owns private $r_\var{priv\_stk}$ region.
Finally, $\rretc$ maps to a world with the ownership of $W$.

we split the world in the same way for the registers that remain unchanged, the
result follows from Lemma \ref{lem:downwards-closed} and \ref{lem:monotonicity}.
This leaves the following cases:
\begin{description}
\item[Case $\rstk$:] \lau{TODO}
\item[Case $\rtmp{1}$:] \lau{TODO}
\item[Case $r_1$,$r_2$:] \lau{TODO}
\item[Case $\rretd$:] We have to show
  \[
    \npair[n'-1]{(\sealed{\sigma',\retptrc(\aaddr_\stk,\eaddr_\stk)},\sealed{\sigma',((\rw,\linear),\aaddr_\stk,\eaddr_\stk,\aaddr_\stk)})} \in \lrv(W_{R,\rretd}')
  \]
  where $W_{R,\rretd}'$ is the part of $W_R'$ with ownership over $r_\var{priv\_stk}$.
  From assumption $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss)$\footnote{This must be the other possible assumption contains $[\baddr,\eaddr] \# \ta$ which leads to a contradition with $[\baddr,\eaddr] \subseteq \ta$}, we know there is a code-region for the

\item[Case $\rretc$:] \lau{TODO}
\end{description}

\begin{lemma}
  If $W \vdash_{(\ta,\stkb)} (\mscode,\_,\_,\_,\sigrets,\sigcloss,\_)$ and
  $[\baddr,\eaddr] \subseteq \dom(\mscode)$ and
  $\exists \aaddr \ldotp \mscode(\aaddr) = \sigma$ and
  $\sigma \in \sigrets$ and
  $\memSat{\dots}{W}$, then
  \[
    \begin{array}{l}
      \exists \mscode' r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) = \codereg{\sigrets',\sigcloss',\mscode',(\ta,\stkb)} \text{ and } \\
      \quad [\baddr,\eaddr] \subseteq \dom(\mscode') \text{ and } \\
      \quad \sigma \in \sigrets' \subseteq \sigrets \text{ and } \\
      \quad \sigclos \subseteq \sigclos'
    \end{array}
  \]
\end{lemma}
